### 2020-10-19 report/draft.tex
fixed brown lines

    Finally, $\delta$ is an extra mortality term defined as
    \[
      \begin{cases}
        d & \mbox{if } t < t_{AE} \\
        0 & \mbox{if } t > t_{AE}
      \end{cases}
    \]
    The parameters values and ranges are reported in \textbf{Table~\ref{tbl1}}.
    %
    \begin{table}
      \centering
      {%
        \begin{tabular}{@{}llccr@{}}
          \toprule
          \textbf{Variable}&\textbf{Meaning} & \textbf{Units} & \textbf{Value} & \textbf{Reference}\\
          \midrule
          ${P}_{0}$       & initial concentration of $P$ cells         & cells/ml          & 4E3         & Experiment detailed in the paper\\
          $\alpha$        & fraction of $P$ that will generate $F_{1}$ & ---               & [0.01, 0.2] & Explored via simulations\\
          $r_{P}$         & net growth rate of $P$                     & $\text{day}^{-1}$ & [1, 3]      & Explored via simulations\\ 
          $r_{F_{1}}$     & net growth rate of $F_{1}$                 & $\text{day}^{-1}$ & 1           & ---\\
          $t_{AE}$        & end day of the growth arrest               & day               & [2,6]       & Explored via simulations\\
          $t_{F_{1}I}$    & day of appearence of the offspring         & day               & 3           & Experiment detailed in the paper\\
          $L_{0,P}$       & starting length for $P$ cells              & $\mu$m            & 40          & Experiment detailed in the paper\\
          $L_{0,{F_{1}}}$ & starting length for $F_{1}$ cells          & $\mu$m            & 80          & Experiment detailed in the paper\\
          $d$             & gametogenesis-induced extra mortality      & day               & [0.1, 0.9]  & Explored via simulations\\
          \bottomrule
        \end{tabular}
       }
      \caption{Values and ranges for model parameters}\label{tbl1}
    \end{table}
%
  \section*{Results and Discussion of the simulations}
    \textbf{Fig.~\ref{fdyn}} shows the relative variation of allometric growth (\textbf{plot {\color{blue}\textit{a}}}) as derived from the equation reported above.
    As the main output variable, we focused on the recruitment success of the offsprings --- quantified as the ratio between $F_{1}$ and $P$ ($r_p : r_{F_{1}}$) --- after ten days from the start of mating.
    Considering the time interval chosen for the simulation (ten days), the size reduction of $P$ and $F_{1}$ (\textbf{plot {\color{blue}\textit{b}}}) is negligible and, with it, the change in the growth rate of $P$ and $F_{1}$, with a significantly slower growth rate for $F_{1}$ due to their larger size.
    \textbf{Plots {\color{blue}\textit{c}}} and {\textbf{\color{blue}{\textit{d}}}} show the variation in abundance of $P$ and $F_{1}$ for two different sets of parameters used for the simulation.
    These plots provide a demonstration of the dynamics that the model simulates.
    A slower growth rate of $P$ ($r_p=1$, \textbf{plot {\color{blue}\textit{d}}} vs $r_p=1.82$, \textbf{plot {\color{blue}\textit{c}}}) that leaves more nitrogen available favours the growth of $F_{1}$, thus producing a greater abundance of $F_{1}$ and a lower abundance of $P$.

    \textbf{\color{brown}
      In the simulations, the whole range of the different parameters with all possible combinations, generating a wide set of scenarios corresponding to the chosen ranges of variation of the different parameters reported in \textbf{Table}~\ref{tbl1}.
    }

    The ten days length for growth arrest was chosen arbitrarily as a realistic time interval during which parental and daughter cells could remain close one to the other in a low sheared flow regime, so that the hypothesis of resources sharing could hold.
    The time spent in close vicinity by cells in a viscous regime  --- as the one we found in situ --- has been seldom analyzed in detail (see Jenkinson \& Wyatt, 1992 [Journal of Plankton Research, 14:1697--1721]).
    However, we assume that to permit mating, which implies direct contact among mates, several cells should converge by some mechanism --- for example, the one hypothesized by Botte et al., 2013 [J. Plankton Research 35: 914--918] and recently supported with observations by Font-Mu{\~n}oz et al.~(2019) [PNAS 116:15997--16002].
    On the basis of this assumtpion, our scenario would simulate local aggregations of cells, whose reproduction is favoured by the reaching of what we define threshold density.
    This concentration should be enough to trigger gametogenesis by chemical signal exchange ({\color{green}\textbf{ref}}).
    Once the gametogenesis starts all the involved cells stop growing and some individuals will form gametes, likely because the triggering caugth them in a specific phase of the cell cycle.
    Only a part of these gametes will encounter the opposite mating type, and only some of these will successfully complete the other phases,producing an $F_{1}$ generation (e.g., Scalco et al.~2014).
    Eventually, those unable to mate or complete gametogenesis will die.
    These considerations are compatible with the negative growth rate of $P$ observed, and are formulated in the model as extra mortality ($d$).

    Assuming that auxospore formation scales with the encounter rate among mates, the number of auxospore formation should positively covary with the number of cell deaths due to unsuccessful encounters.
    This covariation would favour the $F_{1}$ growth, thus favouring better recruitment.

@@

    \begin{itemize}
      \item \textbf{\color{brown} growth ratio is held constant by the large fraction of successful matings and by both the physiological state and cell cycle phase of the mating subpopulation};
      \item \textbf{\color{brown} large fraction of matings depend on then encounter rate};
      \item \textbf{\color{brown} the increase of encounter rate increase $d$};
      \item \textbf{\color{brown} $d$ increase $F_{1}$ success (as shown by the exploration above)}
    \end{itemize}

    We report the values of $\log_{10}(F_{1}/P)$ for different combinations of the parameters as supplementary tables \textbf{Table~\ref{tbl1}}..
    Overall the simulation provides a quantitative assessment of the number of $F_{1}$ that accumulate after ten day from the start of gametogenesis in relation to the number of mating cells and show that:
    1.\ only in extreme scenarios $F_{1}$ abundance will be equal or higher than $P$ and
    2.\ that the biological constrains of gametogenesis (growth arrest and extra death) due to resetting of cell physiology are compensated by an advantage for the $F_{1}$, mimicking a sort of maternal care.

>>>>

    Finally, $\delta$ is an extra mortality term defined as
    \[
      \begin{cases}
        d & \mbox{if } t < t_{AE} \\
        0 & \mbox{if } t > t_{AE}
      \end{cases}
    \]
    The parameters values and ranges are reported in \textbf{Table~\ref{tbl1}}.
    %
    \begin{table}
      \centering
      {%
        \begin{tabular}{@{}llccr@{}}
          \toprule
          \textbf{Variable}&\textbf{Meaning} & \textbf{Units} & \textbf{Value} & \textbf{Reference}\\
          \midrule
          ${P}_{0}$       & initial concentration of $P$ cells         & cells/ml          & 4E3         & Experiment detailed in the paper\\
          $\alpha$        & fraction of $P$ that will generate $F_{1}$ & ---               & [0.01, 0.2] & Explored via simulations\\
          $r_{P}$         & net growth rate of $P$                     & $\text{day}^{-1}$ & [1, 3]      & Explored via simulations\\ 
          $r_{F_{1}}$     & net growth rate of $F_{1}$                 & $\text{day}^{-1}$ & 1           & ---\\
          $t_{AE}$        & end day of the growth arrest               & day               & [2,6]       & Explored via simulations\\
          $t_{F_{1}I}$    & day of appearence of the offspring         & day               & 3           & Experiment detailed in the paper\\
          $L_{0,P}$       & starting length for $P$ cells              & $\mu$m            & 40          & Experiment detailed in the paper\\
          $L_{0,{F_{1}}}$ & starting length for $F_{1}$ cells          & $\mu$m            & 80          & Experiment detailed in the paper\\
          $d$             & gametogenesis-induced extra mortality      & day               & [0.1, 0.9]  & Explored via simulations\\
          \bottomrule
        \end{tabular}
       }
      \caption{Values and ranges for model parameters}\label{tbl1}
    \end{table}
%
  \section*{Results and Discussion of the simulations}
    \textbf{Fig.~\ref{fdyn}} shows the relative variation of allometric growth (\textbf{plot {\color{blue}\textit{a}}}) as derived from the equation reported above.
    Considering the time interval chosen for the simulation (ten days), the size reduction of $P$ and $F_{1}$ (\textbf{plot {\color{blue}\textit{b}}}) is negligible and, with it, the change in the growth rate of $P$ and $F_{1}$, with a significantly slower growth rate for $F_{1}$ due to their larger size.
    \textbf{Plots {\color{blue}\textit{c}}} and {\textbf{\color{blue}{\textit{d}}}} show the variation in abundance of $P$ and $F_{1}$ for two different sets of parameters used for the simulation.
    These plots provide a demonstration of the dynamics that the model simulates.
    A slower growth rate of $P$ ($r_p=1$, \textbf{plot {\color{blue}\textit{d}}} vs $r_p=1.82$, \textbf{plot {\color{blue}\textit{c}}}) that leaves more nitrogen available favours the growth of $F_{1}$, thus producing a greater abundance of $F_{1}$ and a lower abundance of $P$.

    The ten days length for growth arrest was chosen arbitrarily as a realistic time interval during which parental and daughter cells could remain close one to the other in a low sheared flow regime, so that the hypothesis of resources sharing could hold.
    The time spent in close vicinity by cells in a viscous regime  --- as the one we found in situ --- has been seldom analyzed in detail (see Jenkinson \& Wyatt, 1992 [Journal of Plankton Research, 14:1697--1721]).
    However, we assume that to permit mating, which implies direct contact among mates, several cells should converge by some mechanism --- for example, the one hypothesized by Botte et al., 2013 [J. Plankton Research 35: 914--918] and recently supported with observations by Font-Mu{\~n}oz et al.~(2019) [PNAS 116:15997--16002].
    On the basis of this assumtpion, our scenario would simulate local aggregations of cells, whose reproduction is favoured by the reaching of what we define threshold density.
    This concentration should be enough to trigger gametogenesis by chemical signal exchange ({\color{green}\textbf{ref}}).
    Once the gametogenesis starts all the involved cells stop growing and some individuals will form gametes, likely because the triggering caugth them in a specific phase of the cell cycle.
    Only a part of these gametes will encounter the opposite mating type, and only some of these will successfully complete the other phases,producing an $F_{1}$ generation (e.g., Scalco et al.~2014).
    Eventually, those unable to mate or complete gametogenesis will die.
    These considerations are compatible with the observed negative growth rate of $P$, and are formulated in the model as extra mortality ($d$).

    Assuming that auxospore formation scales with the encounter rate among mates, the number of auxospore formation should positively covary with the number of cell deaths due to unsuccessful encounters.
    This covariation would favour the $F_{1}$ growth, thus favouring better recruitment.
    In the following we discuss the main patterns resulting from simulations where the different parameters were varied, in all possible combinations, within the ranges reported in \textbf{Tab.}~\ref{tbl1}
    As the main output variable, we focused on the recruitment success of the offsprings --- quantified as the ratio between $F_{1}$ and $P$ ($r_p : r_{F_{1}}$) --- after ten days from the start of mating.

@@

    We report the values of $\log_{10}(F_{1}/P)$ for different combinations of the parameters as supplementary tables \textbf{Table~\ref{tbl2}}..
    Overall the simulation provides a quantitative assessment of the number of $F_{1}$ that accumulate after ten day from the start of gametogenesis in relation to the number of mating cells and show that:
    1.\ only in extreme scenarios $F_{1}$ abundance will be equal or higher than $P$ and
    2.\ that the biological constrains of gametogenesis (growth arrest and extra death) due to resetting of cell physiology are compensated by an advantage for the $F_{1}$, mimicking a sort of maternal care.
